---
output: html_document
runtime: shiny
---

<link rel="stylesheet" href="styles.css" type="text/css">

<style type="text/css">
body{ /* Normal  */ font-size: 16px; font-family: Arial; color:black}
td {  /* Table  */ font-size: 8px;}
h1.title {font-size: 38px; color: DarkRed;}
h1 { /* Header 1 */ font-size: 28px; color: Black;}
h2 { /* Header 2 */ font-size: 22px; color: Black;}
h3 { /* Header 3 */ font-size: 18px; color: Black;}
code.r{ /* Code block */ font-size: 12px;}
pre { /* Code block - determines code spacing between lines */ font-size: 14px;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Gopher tortoise (*Gopherus polyphemus*) populations experience varying demographic conditions across the species' range in the southeastern United States. This page comprises a flexible tool that allows users to simulate tortoise population growth and dynamics under varying conditions of demographic rates. Specifically, populations experience latitudinal variation in maturity age and fecundity, where more southern populations have faster somatic growth rates, reach sexual maturity at young ages, and lay larger clutches of eggs likely due to increased energy assimilation. To accommodate variation in life history, the user can adjust mean estimates of maturity age, fecundity (clutch size), and survival rates of different life history stages (nests, hatchlings, juveniles, adults). The juvenile stage includes all 1-year old animals up to the year prior to the maturity age. The software flexibly 'unwinds' the demographic rates to appropriate ages and projects the population using an age-based model The model is a female-only model and assumes a pre-breeding census.

See Folt et al. 2022 *Global Ecology and Conservation* for a review of geographic variation of tortoise demographic rates.

```{r pva, echo=FALSE}

library(popbio)
library(mpmtools)

shinyApp(
  
  fluidPage(
    sidebarLayout(
      sidebarPanel(
        h4(p(strong("Demographic Rates"))),
        numericInput("ma", "Age of maturity:", 18, min = 5, max = 30),
        numericInput("bp", "Probability that a female lays eggs:", 0.97, 0, 1, step=0.01),
        numericInput("f", "Clutch size:", 6, 0, 25, step=1),
        numericInput("ns", "Probability of nest survival:", 0.35, 0, 1, step=0.01),
        numericInput("ve", "Probability of egg viability:", 0.85, 0, 1, step=0.01),
        numericInput("pf", "Probability of female:", 0.5, 0, 1, step=0.01),
        numericInput("s_h", "Probability of hatchling survival:", 0.13, 0, 1, step=0.01),
        numericInput("s_j", "Probability of juvenile survival:", 0.75, 0, 1, step=0.01),
        numericInput("s_a", "Probability of adult survival:", 0.96, 0, 1, step=0.01),
        numericInput("n", "Initial population size:", 50, 0, 10000, step=1),
        br(),
        h4(p(strong("Simulation Inputs"))),
        numericInput("nyears", "Projection interval (years)", 50, 1, 100),
        numericInput("nreps", "Number of simulation replicates", 10, 5, 50),
        br(),
        actionButton("run", "Perform simulation",
                     class = "btn-success",
                     style="color: #fff; background-color: #337ab7; border-color: #2e6da4")
        ),

      mainPanel(
        plotOutput("nplot"),
        br(),
        h4(htmlOutput("resultsText")),
        br(),
        div(DT::dataTableOutput("table1"), style = "font-size:75%"),
      )
    )
  ),
  
  server <- function(input, output, session) {
    
      dat <- eventReactive(input$run, {
        
          print(input$run)  # Check if the button click is detected
        
        # Use 'req()' to require certain inputs to perform simulations
        req(input$ma, input$bp, input$f, input$ns, input$ve, input$pf,
            input$s_h, input$s_j, input$s_a,
            # Simulation inputs
            input$nyears, input$nreps)
        
        withProgress(message = "Simulating population growth...", {
        
        library(statmod)
      
        ### 1) Specify User Inputs 
        
        # Population parameters
        ma <- input$ma
        bp <- input$bp
        f <- input$f
        ns <- input$ns
        ve <- input$ve
        pf <- input$pf
        s_h <- input$s_h
        s_j <- input$s_j
        s_a <- input$s_a
        
        # Create a demography schedule, with juvenile and mature age classes
        # The model has a pre-breeding census
        demog_sched <- data.frame(x = 1:ma,
                                  sx = rep(NA, length(1:ma)),
                                  mx = rep(NA, length(1:ma)))
        
        # Specify productivity of juvenile and adult females
        demog_sched[1:(ma-1), "mx"] <- 0
        
        # Specify productivity of juvenile and adult females
        demog_sched[ma, "mx"] <- bp * f * ns * ve * pf *s_h
  
        # Specify juvenile and adult survival rates
        demog_sched[1:(ma-1), "sx"] <- s_j
        demog_sched[ma, "sx"] <- s_a
        
        # Construct a Leslie matrix from this demography schedule
        A <- make_Leslie_matrix(demog_sched)
        
        ## Demographic features
        
        # Calculate the asymptotic growth rate of the population governed by this 
        #   demography schedule:
        lam <- mpmtools::lambda1(A)
        ssd <- popbio::stable.stage(A) # Stable stage (age) distribution
        names(ssd) <- paste0("Age_", 1:length(ssd))
        generation.time <- popbio::generation.time(A) # Generation time
        rv <- popbio::reproductive.value(A)  # Reproductive value
        names(rv) <- paste0("Age_", 1:length(rv))
        elas <- popbio::elasticity(A) # Elasticity values
        elas.values <- round(c(diag(elas[-1,]), elas[ma-1,ma-1], elas[1,ma-1]), 3)
        names(elas.values) <- c(paste0("S_", 1:(length(elas.values)-1)), "Fecundity")
        
        # Save the demographic variables together
        demo_vars <- list(lam, ssd, generation.time, rv, elas.values)
        
        
        ### Project and track population structure
        
        # Starting population size
        n <- input$n
        
        # Simulation parameters
        nyears <- input$nyears
        nreps <- input$nreps
        
        nstages <- dim(A)[1]
        
        # Matrices to save population size and structure
        N_tot <- matrix(0, nyears, nreps)
        N_stages <- array(0, c(nyears, nreps, nstages))
        # array to save population structure across years, reps, and ages
        
        ## Initial population size and structure
        # What is the stable-stage distribution (SSD) of the matrix
        ssd <- stable.stage(A)
        
        # How many individuals in the population?
        n <- 50
        
        # Spread individuals across the SSD
        n_ssd <- ssd * n
        
        # Use poisson draws to randomly populate numbers per stage for each simulation
        #   replicate
        n_i <- matrix(NA, nreps, dim(A)[1])
        for (i in 1:nreps){
          n_i[i,]  <- rpois(length(n_ssd), n_ssd)
        }
        # rowSums(n_i) # usually between 40-60 females to start
        
        ## Partition matrix for simulation
        x <- splitA(A)
        x_T <- x$T
        x_F <- x$F
        
        ## Run the simulation!
        for (j in 1:nreps){ # for each replicate
          n_i_j <- n_i[j,] # specify initial population size, randomly drawn above
          for (i in 1:nyears){ # for each year
            # If it's the first year, specify n_i_j; else,
            #   specify N_stage from previous year
            if(i == 1){
              n_sim <- n_i_j
            } else {
                n_sim <- N_stages[i-1,j,]
                }
            N_stages[i,j,] <- multiresultm(n_sim, x_T, x_F) # project pop
            N_tot[i,j] <- sum(N_stages[i,j,]) # save N
          } #year
        } #rep
        
        ### Save the data so they can be used
        list(N_tot, demo_vars)
        
    }) # end-withProgress
  }) # end-dat
      
  ## Create an output object where we will save all the graphs and tables
  ##  from the Basic Tool
  graphs <- reactiveValues(nplot = NULL, table1 = NULL)

  ## Summary graph 1 - abundance ------------
  output$nplot <- renderPlot({
    
    # Call the data from the simulation run
    N_tot <- dat()[[1]]
    
    # Load ggplot2
    library(ggplot2)
    
    # Convert the matrix to a data frame
    df <- as.data.frame(N_tot)
    
    # Add a column for time (years)
    df$Time <- 1:50  # Assuming 50 years based on the matrix dimensions
    
    # Reshape the data to long format
    df_long <- tidyr::gather(df, key = "Population", value = "Count", -Time)
    
    # Create a ggplot with a logarithmic y-axis
    graphs$nplot <- ggplot(df_long, aes(x = Time, y = Count, color = Population)) +
      geom_line() +
      labs(x = "Time (years)", y = "Total population") +
      ylim(0, max(N_tot)) +
      theme_minimal() +
      theme(legend.position = "none",  # Remove the legend
            axis.title = element_text(size = 18),
            axis.line = element_line(size = 1.25),
            axis.text = element_text(size = 14, color = 1))
           
    
    # Print the graph
    graphs$nplot

  })
  
  ## Summary text
  output$resultsText <- renderUI({
    
    # Recall some parameters
    demo_vars <- dat()[[2]]

    str1 <- paste("Mean population growth rate:", round(demo_vars[[1]], 3))
    str2 <- paste("Generation time (years):", round(demo_vars[[3]], 1))
    HTML(paste(str1, str2, sep='<br/> <br/>'))

  })
  
  },

  options = list(height = 1200)  
)

```